Highly Available Transactions: Virtues and Limitations

### 2013-03-26 writeup/intro.tex
intro tweaks

The relationship of the CAP Theorem and ACID semantics is poorly
understood. While serializability---the gold standard of SSI
transactions---is unavailable, databases have a long tradition of
providing weaker isolation and consistency guarantees. While these
weaker guarantees do not provide SSI for arbitrary general-purpose
read/write transactions, they often admit greater concurrency and
higher performance than comparable SSI algorithms. Accordingly, weak
isolation is widely employed (Section~\ref{needed}), overwhelmingly as
the default setting in today's traditional ``ACID'' and up-and-coming
``NewSQL'' databases and often as the maximum guarantees offered
(Section~\ref{sec:modernacid}). While the prevalence of these
configurations is encouraging, the current literature does not provide
guidance as to the operation of weakly consistent isolation guarantees
in a distributed context. Rather, we have a strong understanding of
weak isolation in the single-server context from which it originated
as well as several techniques for providing distributed
serializability but we have little to no insight into which ACID
guarantees are compatible with Highly Available Transactions (HATs),
or can be provided with high availability and low latency.

This work attempts to bridge the gap between ACID properties and
highly available systems. We analyze a range of existing ACID and
distributed consistency models to provide a rigorous taxonomization of
which models are achievable in a HAT system.  While several
guarantees, like Serializability and Snapshot Isolation, are
unachievable due to reliance on preventing Lost Update and Write Skew
phenomena, many other ACID guarantees are HAT compliant, such as
providing Read Committed isolation, preventing Fuzzy Read phenomena,
and ensuring atomicity of transactional updates. We also attempt to
bridge the often distant worlds of distributed system consistency and
ACID database consistency by analyzing the relationship between
so-called session guarantees, ``register'' semantics, and ACID
properties. In doing so, we introduce a hierarchy of highly available
properties as, for example, some guarantees require ``stickiness,'' or
fate-sharing between clients and database servers. We believe that
this analysis is the first that has been performed over such a broad
range of ACID and distributed consistency models.

>>>>

The relationship of the CAP Theorem and ACID semantics is poorly
understood. While serializability---the gold standard of SSI
transactions---is unavailable, databases have a long tradition of
providing weaker isolation and consistency guarantees. While these
weaker guarantees do not provide SSI for arbitrary general-purpose
read/write transactions, they often admit greater concurrency and
higher performance than comparable SSI algorithms. Accordingly, weak
isolation is widely employed, overwhelmingly as the default setting in
today's traditional ``ACID'' and up-and-coming ``NewSQL'' databases
and often as the maximum guarantees offered
(Section~\ref{sec:modernacid}). While the prevalence of these
configurations is encouraging, the current literature does not provide
guidance as to the operation of weakly consistent isolation guarantees
in a distributed context. Rather, we have a strong understanding of
weak isolation in the single-server context from which it
originated~\cite{adya, ansicritique, gray-isolation} as well as
several techniques for providing distributed
serializability~\cite{bernstein-concurrency} but we have little to no
insight into which ACID guarantees are compatible with Highly
Available Transactions (HATs), or can be provided with high
availability and low latency.

This work attempts to bridge the gap between ACID properties and
highly available systems. We analyze a range of existing ACID and
distributed consistency models to provide a rigorous taxonomization of
which models are achievable in a HAT system.  While several
guarantees, like serializability and Snapshot Isolation, are
unachievable due to reliance on preventing Lost Update and Write Skew
phenomena, many other ACID guarantees are HAT compliant, such as
providing Read Committed isolation, preventing Fuzzy Read phenomena,
and ensuring atomicity of transactional updates. We also attempt to
bridge the often distant worlds of distributed system consistency and
ACID database consistency by analyzing the relationship between
so-called session guarantees~\cite{sessionguarantees}, ``register''
semantics~\cite{herlihy-art}, and ACID properties. In doing so, we
introduce a hierarchy of highly available properties as, for example,
some guarantees require ``stickiness,'' or fate-sharing between
clients and database servers. We believe that this analysis is the
first that has been performed over such a broad range of ACID and
distributed consistency models.
